---
title: "Homework 2"
author: "Matthew Perrotta"
date: "March 19, 2019"
output: pdf_document
---

### Load Packages
```{r echo = T, message = FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(ISLR)
library(glmnet)
library(caret)
library(corrplot)
library(plotmo)
library(boot)
library(pls)
library(mgcv)
```

### Load Data
```{r}
concrete = read.csv('./data/concrete.csv')
```

```{r}
conX = model.matrix(CompressiveStrength~., concrete)

conY = concrete$CompressiveStrength
```

# Q1 Plot
```{r}
featurePlot(conX, conY, plot = 'scatter', labels = c('','Compressive Strength'), type = c('p'))
```

# Q2 Polynomial Regression

### Using Cross-Validation to Determine Degree 'd'
```{r}
set.seed(1)

d = rep(NA, 4)
for (i in 1:4) {
    fit = glm(CompressiveStrength ~ poly(Water, i), data = concrete)
    d[i] = cv.glm(concrete, fit, K = 10)$d[1]
}

plot(1:4, d, xlab = "Degree 'd'", ylab = "MSE", type = "l")
```

By using cross-validation, a degree of 4 is shown to be the best from values ranging from 1 to 4.

### ANOVA
```{r}
fit1 = lm(CompressiveStrength ~ Water, data = concrete)
fit2 = lm(CompressiveStrength ~ poly(Water, 2), data = concrete)
fit3 = lm(CompressiveStrength ~ poly(Water, 3), data = concrete)
fit4 = lm(CompressiveStrength ~ poly(Water, 4), data = concrete)

anova(fit1, fit2, fit3, fit4)
```

Running an ANOVA on 4 models, models 2 through 4 all have very low p values, but model 2 has the lowest p value. 

#Q3 Smoothing Spline
```{r}
fit.ss = smooth.spline(concrete$Water, concrete$CompressiveStrength)
fit.ss$df

waterlims = range(concrete$Water)
water.grid <- seq(from = waterlims[1], to = waterlims[2])

pred.ss = predict(fit.ss,
                   x = water.grid)

pred.ss.df = data.frame(pred = pred.ss$y,
                         water = water.grid)

p = ggplot(data = concrete, aes(x = Water, y = CompressiveStrength)) +
     geom_point(color = rgb(.2, .4, .2, .5))
p +
geom_line(aes(x = water, y = pred), data = pred.ss.df,
          color = rgb(.8, .1, .1, 1)) + theme_bw()
```

```{r}
par(mfrow = c(3,4))
all.dfs = rep(NA, 12)
for (i in 2:13) {
  fit.ss = smooth.spline(concrete$Water, concrete$CompressiveStrength, df = i)
  
  pred.ss = predict(fit.ss, x = water.grid)
  
  plot(concrete$Water, concrete$CompressiveStrength, cex = .5, col = "darkgrey",
       xlab = "Water",
       ylab = "Compression Strength")
  title(paste("Degrees of freedom = ", round(fit.ss$df)),  outer = F)
  lines(water.grid, pred.ss$y, lwd = 2, col = "blue")
}
```

Increasing degrees of freedom increases the flexibility of the curve.

# GAM
```{r}
gam.m1 <- gam(CompressiveStrength ~ Cement + BlastFurnaceSlag + FlyAsh + Water + Superplasticizer + CoarseAggregate + FineAggregate + Age, data = concrete)
gam.m2 <- gam(CompressiveStrength ~ Cement + BlastFurnaceSlag + FlyAsh + s(Water) + Superplasticizer + CoarseAggregate + FineAggregate + Age, data = concrete)

anova(gam.m1, gam.m2, test = "F")

plot(gam.m2)
```

